Bayesian spectral density estimation using P-splines with quantile-based knot placement

### 2020-10-06 responses/responses2/response.tex

\pagestyle{empty}
\begin{document}
\maketitle\thispagestyle{empty}

% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------

\subsection*{Referee \#1}

\response{We would like to thank the referee for a thorough review of our paper. Below, you find a point-by-point response to your points raised. We have now carefully revised the paper taking all of your -- and the other referees' -- comments into account and hope that you will now find the paper acceptable for publication. For your convenience, all changes made are marked in red in the manuscript.  All tables and figures have been modified.}

@@

\begin{enumerate}
\item
\refereeQuote{The authors' terminology is surprising. They compute a P-spline fit to a power spectrum, but they call it a P-spline prior. The P-spline coefficients are the parameters of the model and they have a prior.}\smallskip

\response{We have edited the manuscript and use {\em P-spline prior} only when referring to the prior but use {\em posterior distribution based on the P-spline prior} when referring to the posterior distribution or Bayesian estimate of the spectral density. We use {\em P-spline prior} in analogy to {\em B-spline Dirichlet process prior} and {\em Bernstein-Dirichlet process prior}.  However, the referee raises a subtle point here in what frequentist and Bayesian regard as unknown parameters. In our view, the unknown parameter in this model is the spectral density function. As a prior on this spectral density function, we use a mixture of P-spline distributions. The mixture weights are hyperparameters with their own hyperprior distribution.}\bigskip

\item
\refereeQuote{To get a better fit to sharp peaks in the spectrum, the authors propose an adaptive scheme for knot placement. Apparently, they forgot that the differences in the penalty must change too, to correct for the variations in the distance between knots. One solution could be to use divided differences, another to use the algorithm of Wood for a derivative-based penalty.}\smallskip

\response{Thank you very much for pointing this out. You are absolutely correct, we did use the same penalty for the unequally spaced knots as for the equidistant knots.
We have rectified this and now use derivative-based penalties for the unequally spaced knots. We describe the prior specification in paragraph 4  of the section  ``adaptive knots" in Chapter 3. As third order derivative-based penalties do not make sense for cubic splines and also because we want to compare the results for the equally-spaced knots and adaptive knots in the simulation study, we only compare results based on first and second order difference and derivative-based penalties, respectively. Therefore we deleted the definition of the third order difference matrix in Chapter 3. }\bigskip

@@

\item
\refereeQuote{A more principle approach is to model the logarithm of the spectrum. That can be combined with adaptive smoothing, as described by Wood and Fasiolo, or Rodrigues-Alvarez et al. R packages are available.}\smallskip

\response{It is debatable whether putting a prior on the log spectral density is more {\em principled} than putting a prior on the spectral density itself as both terms enter the Whittle likelihood function (1) and thus the prior can be put on either scale. Our P-spline prior is based on the idea of approximating  a {\em density} function as a mixture of basis distributions on the unit interval.
Thus, the original scale and not the log scale is appropriate for this particular approach even though for many applications, the log-scale might be the more natural one and we cited many papers that put a prior on the log spectral density, see our paragraph 5 of the introduction. This could be an interesting avenue to pursue in future research and see whether putting a P-spline based prior on the log spectral density has advantages over putting a prior on the spectral density. We added a paragraph in the Discussion. We have now also added the references to the alternative frequentist approaches that you suggested. Thank you for pointing these out.

We were wondering whether the referee meant ``model the logarithm of the periodogram". One might consider to model the log-periodogram instead of the periodogram itself which would require using the logarithm of an exponential distribution or approximating
this via a mixture of Gaussian distributions as in Carter and Kohn (1997). However, here we stick to the traditional Whittle likelihood which is contiguous to true likelihood for Gaussian stationary time series (Kirch et al 2019).
}\bigskip

@@

\item
\refereeQuote{The plots of spectra and fit with logarithmic are popular in spectrum analysis, but they are a bit misleading as the fitting is done on the linear scale.}\smallskip

\response{Please see our response to the previous point. Even though we fit a mixture of B-spline densities to the normalized spectral density, it is still possible to show the fit
on the log scale. Large peaks are better visible using the log scale. We also want to have a direct comparison of our model fits in Figure 1 and Figure 2 to the corresponding ones shown  in Edwards et al. (2019) so it makes sense to use the same scale.}\bigskip

\item
\refereeQuote{I took a short look at the repository. It could be made more valuable by adding the data and some example scripts.}\smallskip

\response{Data and example scripts are included in the R package \texttt{psplinePsd}.}\bigskip

@@

\item
\refereeQuote{Sunspots counts are relatively familiar, but very few people will have seen a time the light curve of S. Carinae. I propose to add a plot of (a part of) the time series.}\smallskip

\response{We have added a time series plot of the first 150 observations of the \textit{S.\ Carinae} variable star light intensities data in Figure 4}.\bigskip

\item
\refereeQuote{
The term `P-spaced knots' is confusing in the context of P-splines, P is already used for penalties. An adaptive scheme is proposed that is not exclusive to periodograms. Simply calling them adaptive knot looks adequate to me.}\smallskip

\response{ This could indeed be confusing. We have changed `P-spaced' to `adaptive' knots.}\bigskip

@@

\response{We have deleted the detailed description of the BspDP prior but kept the definition of the B-splines as these are required for the definition of the prior based on the  P-splines.}\bigskip

\item 
\refereeQuote{The proposed method deals with estimating one single spectral density. However, many existing paper already deal with this. Can the proposed methodology be extended to multiple or multivariate time series?}\smallskip

\response{
 This could be a
very interesting avenue for future research and we have added a paragraph in the Discussion suggesting a potential approach. However, a multivariate extension is outside the scope of this paper.}\bigskip

\item
\refereeQuote{A Metropolis step is used for v, but a data augmentation step based on Polya-Gamma auxiliary variables will allow for a full Gibbs sampler. See ``Bayesian Inference for Logistic Models Using PolyaGamma Latent Variables" (2013) of Polson, Scott and Windle.}\smallskip

\response{We had already pointed this out in section ``posterior computation" and had cited the paper by Polson et al. (2013).}

### 2020-10-06 responses/responses3/response.tex

\documentclass{scrartcl}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{color}
\usepackage[english]{babel}
%\usepackage[fixlanguage]{babelbib}
\usepackage{natbib}
\usepackage[pdftex]{graphicx}
\usepackage{dsfont}
\usepackage{color}
\usepackage{hyperref}  % Package for hyperlinking
%\usepackage{a4wide}
\usepackage[a4paper, left=3cm, right=3cm, top=2cm]{geometry}
%\usepackage{texititel}
\title{Bayesian spectral density estimation using P-splines with quantile-based knot placement}
\subtitle{Response letter to Statistical Papers}
\author{Patricio Maturana-Russel, Renate Meyer}

\newcommand{\refereeQuote}{\textit }
\newcommand{\responseDraft}{\textcolor{red}}
\newcommand{\response}{}

\parindent 0pt
\parskip 4pt

\setlength{\oddsidemargin}{0cm}
\setlength{\evensidemargin}{0cm}
\setlength{\topmargin}{-2cm}

\pagestyle{empty}
\begin{document}
\maketitle\thispagestyle{empty}

% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------

\subsection*{Referee \#1}

\response{We would like to thank the referee for a thorough review of our paper. Below, you find a point-by-point response to your points raised. We have now carefully revised the paper taking all of your -- and the other referees' -- comments into account and hope that you will now find the paper acceptable for publication. For your convenience, all changes made are marked in red in the manuscript.}

\textbf{Major comment:}

\refereeQuote{p.8, Quantile-based knot placement\\
I could not understand the meaning of ``calculate its distribution function $F_z = P(z \leq z_i)$" in the second item of the procedure.  The values of the transformed periodogram $z_i$ are regarded as the values of the density $\tilde{f}$, which is non-monotonic, taken at equally-spaced frequencies.  According to item 4, the distribution function $F_z$, and hence the density function $\tilde{f}$, should be defined on the frequency domain. \\
I had a look at the R codes of the package psplinePsd. I think that the density function $\tilde{f}$ is a linear spline that interpolates as
\begin{align*}
\tilde{f} \left(\dfrac{\lambda_k}{2\pi}\right) = \dfrac{z_k}{\sum_{i=1}^{\nu}z_i}, \quad k=1,\dots,\nu.
\end{align*}
}

\response{That's correct!  The $\widetilde{f}$ is actually a probability mass function (not a probability density function), which is interpolated.  We have corrected this in the text.  We have also changed the description of the algorithm, following now the exact steps implemented in the \texttt{psplinePsd} package.}\bigskip

\textbf{Minor comments:}
\begin{enumerate}
\item
\refereeQuote{p.4, line 10\\
A spline of degree $r+1\implies$ A spline of degree $r$ (or, of order $r+1$)}\smallskip

\response{Fixed.}\bigskip

\item
\refereeQuote{p.5 line 13\\
In the definition of $\tau, f(\tau \omega)\implies f(\pi \omega)$}\smallskip

\response{That is correct.  Fixed. }\bigskip

\end{enumerate}

% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------

\subsection*{Referee \#2}

\response{We would like to thank the referee for a thorough review of our paper. Below, you find a point-by-point response to your points raised. We have now carefully revised the paper taking all of your -- and the other referees' -- comments into account and hope that you will now find the paper acceptable for publication. For your convenience changes are marked in red in the manuscript.
} \smallskip

\textbf{Specific comments:}
\begin{enumerate}
	\item \refereeQuote{ Page 5, how to choose penalty or prior $\phi$?}

	\response{In a Bayesian setting, the roughness penalty is translated into the prior distribution for the spline parameters via a normal distribution.  See for instance \cite{Bremhorst:2016}.  In order to make the posterior less sensitive to the prior, we suggest to follow their proposal and use their robust specifications for the hierarchical priors on $\phi$, which is also described in our paper. }

	\item \refereeQuote{To select the number of B-spline densities, the authors employ a heuristic rule of thumb $K^*$ in the application Section, it is of interest to compare it with other criteria such as AIC and BIC.}\smallskip

	\response{This could be a very interesting avenue for future research, however, we think it is beyond the scope of this paper.  An analysis of those characteristics can be quite extensive and could constitute a paper in itself, see for instance \cite{Likhachev2017}.  We would like to highlight that the number of knots is considered non-important by many, since the penalty parameter plus the penalty order difference penalties specified in the penalty matrix $\textbf{P}$ penalise the inclusion of B-spline densities, controlling the smoothness of the estimate.  This means that we only need to select a minimum value, the penalty terms take care of the rest.}\bigskip

	Other minor comments: \bigskip

	\item \refereeQuote{Page 2, line 17, what is $\lambda_i$?}\smallskip

	\response{$\lambda_i$ represents any angular frequency. Since we only defined the angular frequency $\lambda$ in that paragraph, we deleted its subscript ``$i$" and added $\lambda\in[-\pi,\pi]$. } \bigskip

	\item \refereeQuote{ Page 4, lines 5-7, ``In particular, the approximation error of the mixture of B-spline distributions can be made arbitrarily small by finding suitable knot locations and increasing the number of knots", what does ``made arbitrarily small" mean? Do you mean the approach with the beta densities the error cannot be made arbitrarily small?}\smallskip

	\response{Yes, the loss associated to the approximation given by the beta densities with respect to a loss function cannot be made arbitrarily small. This was shown by \cite{Perron:2001} who also showed that if one replaces the beta densities by B-spline densities of fixed order but with variable knots, the loss can be made arbitrarily small by increasing the number of knots.  We have added the corresponding reference to our manuscript.} \bigskip

	\item \refereeQuote{Page 4, line 15, ``A spline of degree r + 1 is a function", I think degree" should be order", we use degree for polynomials.}\smallskip

	\response{Fixed.} \bigskip

	\item \refereeQuote{Page 5, line 20, ``If $\tau = \int_0^1f(\tau \omega)d \omega$ is the normalizing constant", it implies $s_r$ is a density, is it correct?}

	\response{Yes, $s_r$ is a pdf.  Note that it is actually a mixture of B-spline densities.}
\end{enumerate}

\subsection*{Further Changes}
Here some extra changes:
\begin{enumerate}
	\item We have specified the prior on $\tau$. 
	\item We have added $s_r(\omega)$ in Eq(3), short notation used in the posterior distribution of~$\tau$.	
\end{enumerate}

\bibliographystyle{plainnat}
\bibliography{pSplinesReferences}  % Name of BibTeX file

%\begin{thebibliography}{38}
%	\providecommand{\natexlab}[1]{#1}
%	\providecommand{\url}[1]{\texttt{#1}}
%	\expandafter\ifx\csname urlstyle\endcsname\relax
%	\providecommand{\doi}[1]{doi: #1}\else
%	\providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
%	\bibitem[Bremhorst and Lambert(2016)]{Bremhorst:2016}
%	Vincent Bremhorst and Philippe Lambert.
%	\newblock Flexible estimation in cure survival models using {B}ayesian
%	p-splines.
%	\newblock \emph{Computational Statistics \& Data Analysis}, 93:\penalty0 270 --
%	284, 2016.
%	\newblock ISSN 0167-9473.
%	\newblock \doi{https://doi.org/10.1016/j.csda.2014.05.009}.
%	\newblock URL
%	\url{http://www.sciencedirect.com/science/article/pii/S0167947314001492}.
%\end{thebibliography}

\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-10-06 pSplines.tex
revision

\newcommand{\pmr}{ \color{blue}}

@@

As the periodogram fluctuates around the true spectral density, most  nonparametric methods for spectral density estimation hinge on  smoothing the periodogram
\begin{align*}
I_n(\lambda) = \dfrac{1}{2 \pi n} \left| \sum_{t=1}^{n} Y_t \exp \left( -i t \lambda\right)\right|^2,
\end{align*}
in one way or another, where $-\pi < \lambda \leq \pi$ is the angular frequency and $\bm{Y}=(Y_1,\ldots,Y_n)$ is a stationary time series of length $n$. The periodogram ordinates evaluated at the Fourier frequencies $\lambda_i$  are asymptotically independent and follow an exponential distribution with mean $f(\lambda_i)$.  Thus, they are unbiased estimates of $f(\lambda_i)$  but not consistent \citep{Brockwell:1986}.

Most Bayesian approaches to psd estimation use the so-called Whittle likelihood, introduced by \cite{Whittle:1957} as an approximation to the likelihood of a Gaussian stationary time series.  Based on the  limiting independent exponential distribution of the periodogram values, the Whittle likelihood for a mean centred weakly stationary time series $\bm{Y}$ of length $n$ is defined as
\begin{align}
L(\bm{Y} | f) &= \prod_{l=1}^{\nu}\dfrac{1}{f(\lambda_{l})} e^{-I_n(\lambda_l)/f(\lambda_l)}  \nonumber \\
&\propto \exp \left( - \sum_{l=1}^{\nu} \log f(\lambda_l) + \dfrac{I_{n}(\lambda_l)}{f(\lambda_l)} \right),\label{eq:whittle_like}
\end{align}
where $\lambda_l = 2\pi l / n$ are the Fourier frequencies, $f(\cdot)$ spectral density of $\bm{Y}$, and $\nu = \left\lfloor (n-1)/2 \right\rfloor$ is the greatest integer value less than or equal to $(n-1)/2$. Advantages of this approximation are that it depends directly on the spectral density, unlike the true Gaussian likelihood function, and gives a good approximation for non-Gaussian time series under certain conditions as shown by \cite{ShaoXiaofeng2007ASTf}.

@@

As shown by \cite{Perron:2001} for distributions on the interval [0,1], the mixture of B-spline distributions has better approximation properties than the mixture of beta distributions for a fixed number of components. In particular, the approximation error of the mixture of B-spline distributions 
can be made arbitrarily small by finding  suitable knot locations and increasing the number of knots.
%the mixture of beta distributions is unable to provide a full coverage of the space of cdfs on $[0,1]$.  They defined a loss function and showed that this cannot be arbitrarily small under this %approach.  Alternatively, they showed that if the beta distributions are replaced by B-spline distributions of fixed order (shown for order~2) with variable knots, the loss function can be %arbitrarily small by increasing the number of knots.  The adequacy of the approximation is a function of the number of knots.  
This work motivated \cite{Edwards2019} to propose the BspDP prior on the spectral density function which is based on the representation (\ref{eq:Bernstein}) but with
the beta densities replaced by B-spline densities as defined in the following.

@@

A spline of degree $r+1$ is a function defined piecewise by polynomials of degree $\leq r$, which meet at points called \textit{knots} where the function is continuous.  Any of these spline functions can be described by basis functions, known as \textit{B-splines}, in other words, all spline functions can be represented as a unique combination of B-splines with the same order and over the same partition.  Without loss of generality, the global domain of interest is assumed to be the unit interval $[0,1]$.

@@

As the spectral density function $f(\cdot)$ is defined on the interval $[0,\pi]$, it is reparametrized and defined as
\begin{align*}
f(\pi \omega) =\tau \times s_r(\omega;\bm{w}, \bm{\xi}), \quad \omega \in [0,1], %\tau \times s_r(\omega;K,G,H), \quad \omega \in [0,1],
\end{align*}
where $\tau = \int_{0}^{1}f(\tau \omega)\text{d}\omega$ is the normalizing constant and
\begin{align}
s_r(\omega; \bm{w}, \bm{\xi}) = \sum_{k=1}^{K} w_k b_{k,r}(\omega;\bm{\xi}),
\label{eq:ps_prior}	
\end{align}
with  {\em fixed} number $K$ of B-spline densities  $b_{k,r}(\cdot)$ of fixed degree $r$, weight vector $\bm{w}=(w_1,\cdots, w_K)$, and knot sequence $\bm{\xi}$.

@@

The parameter vector for the P-spline prior is $\bm{\theta} = (\bm{v}^\top, \phi, \delta, \tau)^\top$.
We use a robust specification for the  prior distribution of the penalty parameters as suggested by \cite{Jullion:2007} by choosing small values for $\alpha_{\delta}$ and $\beta_{\delta}$, for instance, $10^{-4}$. The  choice of $\alpha_{\phi}$ and $\beta_{\phi}$ do not affect the spectral density estimate. Here we set $\alpha_{\phi}$ and $\beta_{\phi}$ equal to 1 as used by \cite{Bremhorst:2016}. The scale parameter $\tau$ is given by an inverse Gamma($\alpha_{\tau},\beta_{\tau}$) prior.  These are also the prior specifications used in the simulations and examples in Section \ref{sec:application}.

@@

Sampling from the full conditional posterior distributions of $\phi$, $\delta$, and $\tau$ can be performed directly. These are given by
%The conditional posterior of $\phi$, $\delta$, and $\tau$ belong to known families of distributions, that is
\begin{align*}
\phi|\bm{Y},\delta,\bm{v} &\sim \text{Gamma}\left(\tfrac{K-1}{2} + \alpha_{\phi},\: \tfrac{1}{2} \bm{v}^\top \textbf{P} \bm{v} + \delta \beta_{\phi} \right), \\
\delta|\bm{Y},\phi  &\sim \text{Gamma}\left(\alpha_{\phi} + \alpha_{\delta}, \beta_{\phi} \phi + \beta_{\delta}\right),\:\: \text{and}\\
\tau|\bm{Y},\delta, \phi,\bm{v}   &\sim \text{Inv-Gamma}\!\left(\alpha_{\tau}+ \nu,\:  {\textstyle\sum\limits_{l=1}^{\nu}} \tfrac{I_n(\lambda_l)}{f(\lambda_l)} + \beta_{\tau} \right)\!.
\end{align*}
%The posterior distribution of $\tau$ is the same one obtained under the B-spline prior approach.

However, to sample from the full conditional distribution of $\bm{v}$, we use the Metropolis algorithm.
\textcolor{red}{Since the weights in general are larger in those areas of the frequency domain where the psd has peaks, we specify their starting value proportionally to the periodogram. Applying the corresponding transformation we obtain the starting value for the vector $\bm{v}$.  This strategy speeds up the MCMC process by shorting the burn-in period.}

@@

	\item Draw $\delta^{m}$ from $\text{Gamma}\left(\alpha_{\phi} + \alpha_{\delta}, \beta_{\phi} \phi^{m} + \beta_{\delta}\right);$ 
	\item Draw $\tau^{m}$ from 
	\begin{align*}
	\text{Inv-Gamma}\!\left(\alpha_{\tau}\!+\!\nu, {\textstyle\sum\limits_{l=1}^{\nu}} \tfrac{I_n(\lambda_l)}{f(\lambda_l)} + \beta_{\tau} \right) .
	\end{align*}
\end{itemize}

@@

Our knot location proposal works as follow.  We calculate the periodogram and apply a square root transformation in order to have more regular magnitudes and eliminate a potential trend.  Second, we standardize these values and apply the absolute value function.  We treat this transformed periodogram as a density function $\widetilde{f}$ and calculate its distribution function $F$.  Finally, the knots are allocated according to the quantiles of the inverse distribution function of $\widetilde{f}$.  Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes. Hereafter, we will refer to this scheme as {\em  quantile-spaced (Q-spaced) knots}.%\textit{periodogram-spaced knots} or short \textit{P-spaced knots}.
%REMARK: HOW ABOUT "PERIODOGRAM-SPACED KNOTS"	OR 'P-SPACED KNOTS'  INSTEAD OF DISTRIBUTED KNOTS?

The procedure can be summarized as follows:
\begin{enumerate}
	\item Let $\bm{x} = \left (\sqrt{I_n(\lambda_1)}, \dots, \sqrt{I_n(\lambda_{\nu})} \right)$ be the square root of the periodogram values for each angular frequency.
	\item Define $z_i = \left|\dfrac{x_i - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$ and calculate its distribution function $F_z = P(z \leq z_i)$.
	\item Let $\bm{\xi}^* = \left(\xi_1^*,\dots,\xi_{K^*}^* \right)$ be a vector of $K^*$ quantiles, equally spaced in $[0,1]$.
	\item The Q-spaced knot vector is given by $\bm{\xi} = \left(F_{z}^{-1}(\xi^*_1),\dots, F_{z}^{-1}(\xi^*_{K^*}) \right)$.

\end{enumerate}

>>>>

\newcommand{\pmr}{ \color{red}}

@@

As the periodogram fluctuates around the true spectral density, most  nonparametric methods for spectral density estimation hinge on  smoothing the periodogram
\begin{align*}
I_n(\lambda) = \dfrac{1}{2 \pi n} \left| \sum_{t=1}^{n} Y_t \exp \left( -i t \lambda\right)\right|^2,
\end{align*}
in one way or another, where $-\pi < \lambda \leq \pi$ is the angular frequency and $\bm{Y}=(Y_1,\ldots,Y_n)$ is a stationary time series of length $n$. {\pmr The periodogram ordinates evaluated at the Fourier frequencies $\lambda \in [-\pi,\pi]$  are asymptotically independent and follow exponential distributions with mean $f(\lambda)$.  Thus, they are unbiased estimates of $f(\lambda)$  but not consistent \citep{Brockwell:1986}.}

Most Bayesian approaches to psd estimation use the so-called Whittle likelihood, introduced by \cite{Whittle:1957} as an approximation to the likelihood of a Gaussian stationary time series.  Based on the  limiting independent exponential distribution of the periodogram values, the Whittle likelihood for a mean centred weakly stationary time series $\bm{Y}$ of length $n$ is defined as
\begin{align}
L(\bm{Y} | f) &= \prod_{l=1}^{\nu}\dfrac{1}{f(\lambda_{l})} e^{-I_n(\lambda_l)/f(\lambda_l)}  \nonumber \\
&\propto \exp \left\{ -\! \sum_{l=1}^{\nu} \Bigg( \log f(\lambda_l) + \dfrac{I_{n}(\lambda_l)}{f(\lambda_l)} \Bigg) \right\},\label{eq:whittle_like}
\end{align}
where $\lambda_l = 2\pi l / n$ are the Fourier frequencies, $f(\cdot)$ spectral density of $\bm{Y}$, and $\nu = \left\lfloor (n-1)/2 \right\rfloor$ is the greatest integer value less than or equal to $(n-1)/2$. Advantages of this approximation are that it depends directly on the spectral density, unlike the true Gaussian likelihood function, and gives a good approximation for non-Gaussian time series under certain conditions as shown by \cite{ShaoXiaofeng2007ASTf}.

@@

As shown by \cite{Perron:2001} for distributions on the interval [0,1], the mixture of B-spline distributions has better approximation properties than the mixture of beta distributions for a fixed number of components. In particular, the approximation error of the mixture of B-spline distributions can be made arbitrarily small by finding  suitable knot locations and increasing the number of knots {\pmr\citep{Perron:2001}}.
%the mixture of beta distributions is unable to provide a full coverage of the space of cdfs on $[0,1]$.  They defined a loss function and showed that this cannot be arbitrarily small under this %approach.  Alternatively, they showed that if the beta distributions are replaced by B-spline distributions of fixed order (shown for order~2) with variable knots, the loss function can be %arbitrarily small by increasing the number of knots.  The adequacy of the approximation is a function of the number of knots.  
This work motivated \cite{Edwards2019} to propose the BspDP prior on the spectral density function which is based on the representation (\ref{eq:Bernstein}) but with
the beta densities replaced by B-spline densities as defined in the following.

@@

A spline of {\pmr order} $r+1$ is a function defined piecewise by polynomials of degree $\leq r$, which meet at points called \textit{knots} where the function is continuous.  Any of these spline functions can be described by basis functions, known as \textit{B-splines}, in other words, all spline functions can be represented as a unique combination of B-splines with the same order and over the same partition.  Without loss of generality, the global domain of interest is assumed to be the unit interval $[0,1]$.

@@

As the spectral density function $f(\cdot)$ is defined on the interval $[0,\pi]$, it is reparametrized and defined as
\begin{align*}
f(\pi \omega) =\tau \times s_r(\omega;\bm{w}, \bm{\xi}), \quad \omega \in [0,1], %\tau \times s_r(\omega;K,G,H), \quad \omega \in [0,1],
\end{align*}
where $\tau = \int_{0}^{1}f(\pi \omega)\text{d}\omega$ is the normalizing constant and
\begin{align}
s_r(\omega; \bm{w}, \bm{\xi}) = {\pmr s_r(\omega)} = \sum_{k=1}^{K} w_k b_{k,r}(\omega;\bm{\xi}),
\label{eq:ps_prior}	
\end{align}
with  {\em fixed} number $K$ of B-spline densities  $b_{k,r}(\cdot)$ of fixed degree $r$, weight vector $\bm{w}=(w_1,\cdots, w_K)$, and knot sequence $\bm{\xi}$.

@@

The parameter vector for the P-spline prior is $\bm{\theta} = (\bm{v}^\top, \phi, \delta, \tau)^\top$.
We use a robust specification for the  prior distribution of the penalty parameters as suggested by \cite{Jullion:2007} by choosing small values for $\alpha_{\delta}$ and $\beta_{\delta}$, for instance, $10^{-4}$. The  choice of $\alpha_{\phi}$ and $\beta_{\phi}$ do not affect the spectral density estimate. Here we set $\alpha_{\phi}$ and $\beta_{\phi}$ equal to 1 as used by \cite{Bremhorst:2016}. The scale parameter $\tau$ is given by an inverse Gamma($\alpha_{\tau},\beta_{\tau}$) prior, {\pmr which can be considered as a noninformative prior for $\alpha_{\tau} = \beta_{\tau} = 0.001$ \citep{Edwards2019}.}  These are also the prior specifications used in the simulations and examples in Section \ref{sec:application}.

@@

Sampling from the full conditional posterior distributions of $\phi$, $\delta$, and $\tau$ can be performed directly. These are given by
%The conditional posterior of $\phi$, $\delta$, and $\tau$ belong to known families of distributions, that is
\begin{align*}
\phi|\bm{Y},\delta,\bm{v} &\sim \text{Gamma}\left(\tfrac{K-1}{2} + \alpha_{\phi},\: \tfrac{1}{2} \bm{v}^\top \textbf{P} \bm{v} + \delta \beta_{\phi} \right), \\
\delta|\bm{Y},\phi  &\sim \text{Gamma}\left(\alpha_{\phi} + \alpha_{\delta}, \beta_{\phi} \phi + \beta_{\delta}\right),\:\: \text{and}\\
\tau|\bm{Y},\delta, \phi,\bm{v}   &\sim \text{Inv-Gamma}\!\left(\alpha_{\tau}+ \nu,\:  {\textstyle\sum\limits_{l=1}^{\nu}} \tfrac{I_n(\lambda_l)}{{\pmr s_r(\lambda_l/\pi)}} + \beta_{\tau} \right)\!.
\end{align*}
%The posterior distribution of $\tau$ is the same one obtained under the B-spline prior approach.

However, to sample from the full conditional distribution of $\bm{v}$, we use the Metropolis algorithm.
Since the weights in general are larger in those areas of the frequency domain where the psd has peaks, we specify their starting value proportionally to the periodogram. Applying the corresponding transformation we obtain the starting value for the vector $\bm{v}$.  This strategy speeds up the MCMC process by shortening the burn-in period.

@@

	\item Draw $\delta^{m}$ from $\text{Gamma}\left(\alpha_{\phi} + \alpha_{\delta}, \beta_{\phi} \phi^{m} + \beta_{\delta}\right);$ 
	\item Draw $\tau^{m}$ from 
	\begin{align*}
	\text{Inv-Gamma}\!\left(\alpha_{\tau}\!+\!\nu, {\textstyle\sum\limits_{l=1}^{\nu}} \tfrac{I_n(\lambda_l)}{{\pmr s_r(\lambda_l/\pi)}} + \beta_{\tau} \right) .
	\end{align*}
\end{itemize}

@@

Our knot location proposal works as follow.  We calculate the periodogram and apply a square root transformation in order to have more regular magnitudes and eliminate a potential trend.  Second, we standardize these values, apply the absolute value function, {\pmr and normalise them}.  We treat this transformed periodogram as a {\pmr mass probability} function (pmf) $\widetilde{f}$,{\pmr which is then linearly interpolated in order to define a pmf in a more refined grid.}  Then its distribution function $\widetilde{F}$ is {\pmr approximated by interpolation}.  Finally, the knots are allocated according to the quantiles of the inverse distribution function $\widetilde{F}^{-1}$, {\pmr which is approximated by interpolation.}  Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes. Hereafter, we will refer to this scheme as {\em  quantile-spaced (Q-spaced) knots}.%\textit{periodogram-spaced knots} or short \textit{P-spaced knots}.
%REMARK: HOW ABOUT "PERIODOGRAM-SPACED KNOTS"	OR 'P-SPACED KNOTS'  INSTEAD OF DISTRIBUTED KNOTS?

%The procedure can be summarized as follows:
%\begin{enumerate}
%	\item Let $\bm{x} = \left (\sqrt{I_n(\lambda_1)}, \dots, \sqrt{I_n(\lambda_{\nu})} \right)$ be the square root of the periodogram values for each angular frequency.
%	\item Define $z_i = \left|\dfrac{x_i - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$ and calculate its distribution function $F_z = P(z \leq z_i)$.
%	\item Let $\bm{\xi}^* = \left(\xi_1^*,\dots,\xi_{K^*}^* \right)$ be a vector of $K^*$ quantiles, equally spaced in $[0,1]$.
%	\item The Q-spaced knot vector is given by $\bm{\xi} = \left(F_{z}^{-1}(\xi^*_1),\dots, F_{z}^{-1}(\xi^*_{K^*}) \right)$.	
%\end{enumerate}

{\pmr
The procedure for the frequencies in the unit interval $\omega_l = 2 l/n  \in[0,1]$, which can be easily translated to the frequency domain, can be summarized as follows:	
\begin{enumerate}
	\item Let $\bm{x} = \left (\sqrt{I_n(\omega_1)}, \dots, \sqrt{I_n(\omega_{\nu})} \right)$ be the square root of the periodogram values.
	\item Define $y_l = \left|\dfrac{x_l - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$.
	\item Define $z_l = \frac{y_l}{\sum_{l=1}^{\nu} y_l}$, normalised $y_l$ values, and calculate $\widetilde{f}$ by interpolating these values.
	\item For a large sequence of values $\bm{\varepsilon}\in[0,1]$, calculate $\widetilde{f}(\bm{\varepsilon})$ and normalise them.
	\item Interpolate the cumulative normalised $\widetilde{f}(\bm{\varepsilon})$ values, which will produce an approximated distribution function $\widetilde{F}$.	
	\item Let $\bm{\xi}^* = \left(\xi_1^*,\dots,\xi_{K^*}^* \right)$ be a vector of $K^*$ quantiles, equally spaced in $[0,1]$.
	\item The Q-spaced knot vector is given by $\bm{\xi} = \left(\widetilde{F}^{-1}(\xi^*_1),\dots, \widetilde{F}^{-1}(\xi^*_{K^*}) \right)$, where the inverse function is also obtained by interpolation.	
\end{enumerate}

Note that this knot allocation algorithm is applied only once at the beginning of the MCMC process.  Therefore, its computational cost is negligible.

}
